\noindent\rule{13cm}{1.0pt}\\ \vspace{0.2cm} A Polynomial-time Reduction of the 3-SAT to the Quadratic Congruence and other Related Problems \noindent\rule{13cm}{0.8pt}

### 2017-11-18 paper/paper.tex
First draft complete.

\newcommand{\algorithmautorefname}{Algorithm}
\algrenewcommand\algorithmicrequire{\textbf{Input}}
\algrenewcommand\algorithmicensure{\textbf{Output}}
\algrenewcomment[1]{\hspace{0.25cm}\(\triangleright\) #1}
\algnewcommand{\LineComment}[1]{\State\,\(\triangleright\) #1}

@@

\title{%
  \noindent\rule{13cm}{1.0pt}\\
  \vspace{0.2cm}
  A Polynomial-time Reduction of the 3-SAT to the Quadratic Congruence Problem and other Related Problems
  \noindent\rule{13cm}{0.8pt}
}
\xdef\shorttitle{Quadratic congruence polynomial-time reduction}
\author[]{\normalsize Renato Lui Geh\\\small NUSP\@: 8536030\\\\Computational Number Theory
(MAC6927)\\Prof\@. Sinai Robins\\University of São Paulo\\}

@@

German mathematician David Hilbert published~\cite{hilbert} in 1902 a set of 23 unsolved problems
in mathematics he deemed to be most important unanswered mathematical problems of the 20th century.
Since then 9 of them have been solved (at the time the author is writing this line and as far as
the author is aware), 9 are considered partially solved, three of them are unsolved and two of them
are considered too vague. Unsolved problems include the infamous Riemann Hypothesis and an
extension to the Kronecker-Weber Theorem. Amongst solved problems is the 10th Hilbert problem.

@@

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} vector sorting}
  \begin{algorithmic}[1]
    \Require\, $n\in\mathbb{N}$ and a vector $A[1..n]$ with $n$ integers
    \Ensure\, An ordered vector
    \Statex\, Sorting the vector $A$ in increasing order.
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Search problem:} greatest common divisor}
  \begin{algorithmic}[1]
    \Require\,$a, b\neq 0\in\mathbb{N}$
    \Ensure\, $d|a$, $d|b$, and for all $d'|a$, $d'|b$ we have $d'\leq d$
    \Statex\, Finding the $\gcd(a, b)$
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Optimization problem:} longest common subsequence}
  \begin{algorithmic}[1]
    \Require\, Two strings $X[1..m]$ and $Y[1..n]$
    \Ensure\, A string
    \Statex\, Finding the longest common subsequence in $X$ and $Y$.
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Decision problem:} greatest common divisor}
  \begin{algorithmic}[1]
    \Require\,$a, b\neq 0\in\mathbb{N}$ and $k\in\mathbb{N}$
    \Ensure\, Boolean value
    \Statex\, Is $\gcd(a, b)=k$ true?
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
  \caption*{\textbf{Decision problem:} longest common subsequence}
  \begin{algorithmic}[1]
    \Require\, Two strings $X[1..m]$ and $Y[1..n]$
    \Ensure\, Boolean value
    \Statex\, Is there an LCS of X and Y such that its length $\geq k$?
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} Hamiltonian cycle}
  \begin{algorithmic}[1]
    \Require\, A graph $G=(V, E)$
    \Ensure\, Boolean value
    \Statex\, Is there a Hamiltonian cycle, i.e.\ a path $p=(e_1,e_2,\ldots,e_k)$ s.t. $\forall v
    \in V$, $v$ is visited by $p$ exactly once, and $p$ is a cycle?
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} SAT}
  \begin{algorithmic}[1]
    \Require\, A Boolean formula with $n$ variables written in CNF, e.g.\ $(x_1\vee x_2\vee x_2\vee
    x_3)\wedge (\ov{x}_2\vee x_5)\wedge\cdots$
    \Ensure\, Boolean value
    \Statex\, Is there a valuation $\nu :\{x_1,x_2,\ldots,x_n\}\to\{\false, \true\}$ such that the
    input formula evaluates to true?
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} 3-SAT}
  \begin{algorithmic}[1]
    \Require\, A Boolean formula with $n$ variables written in CNF with at most 3 literals in each
    clause, e.g.\ $(x_1\vee x_2\vee x_3)\wedge(\ov{x}_2\vee\ov{x}_1\vee x_5) \wedge\cdots$
    \Ensure\, Boolean value
    \Statex\, Is there a valuation $\nu :\{x_1,x_2,\ldots,x_n\}\to\{\false, \true\}$ such that the
    input formula evaluates to true?
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} quadratic congruence}
  \begin{algorithmic}[1]
    \Require\, $\alpha,\beta,\gamma\in\mathbb{Z}$
    \Ensure\, Boolean value
    \Statex\, Is there a solution $x\in\mathbb{Z}$ where
    \begin{equation*}
      x^2 \equiv\alpha\mod\beta
    \end{equation*}
    and such that $0\leq x\leq\gamma$?
  \end{algorithmic}
\end{algorithm}

@@

\begin{align*}
  (i) \quad &0\leq x_1\leq H, \quad x_1\in\mathbb{Z},\\
  (ii) \quad &\lambda_1 2\cdot 8^{m+1}(H^2-x_1^2)+\lambda_2 K(\tau^2-x_1^2)\equiv 0\mod 2\cdot
    8^{m+1}\cdot K, &\text{(III)}\\
  (iii) \quad &\gcd(\lambda_1,K)=\gcd(\lambda_2,2\cdot 8^{m+1})=1,\quad\lambda_1,\lambda_2\in
    \mathbb{Z}
\end{align*}

@@

%--------------------------------------------------------------------------------------------------

>>>>

\newcommand{\algorithmautorefname}{Algorithm}
\algrenewcommand\algorithmicrequire{\textbf{Input}}
\algrenewcommand\algorithmicensure{\textbf{Output}}

\algnewcommand\algorithmicref{\textbf{Reference}}
\algnewcommand\Reference{\item[\algorithmicref]}
\algnewcommand\algorithmicred{\textbf{Reduction}}
\algnewcommand\Reduction{\item[\algorithmicred]}
\algnewcommand\algorithmicdesc{\textbf{Description}}
\algnewcommand\Description{\item[\algorithmicdesc]}
\algnewcommand\algorithmiccom{\textbf{Comment}}
\algnewcommand\Commentx{\item[\algorithmiccom]}

\algrenewcomment[1]{\hspace{0.25cm}\(\triangleright\) #1}
\algnewcommand{\LineComment}[1]{\State\,\(\triangleright\) #1}

@@

\title{%
  \noindent\rule{13cm}{1.0pt}\\
  \vspace{0.2cm}
  A Polynomial-time Reduction of the 3-SAT to the Quadratic Congruence and other Related Problems
  \noindent\rule{13cm}{0.8pt}
}
\xdef\shorttitle{Quadratic congruence polynomial-time reduction}
\author[]{\normalsize Renato Lui Geh\\\small NUSP\@: 8536030\\\\Computational Number Theory
(MAC6927)\\Prof\@. Sinai Robins\\University of São Paulo\\}

@@

German mathematician David Hilbert published~\cite{hilbert} in 1902 a set of 23 unsolved problems
in mathematics he deemed to be the most important mathematical problems of the 20th century.  Since
then 9 of them have been solved (at the time the author is writing this line and as far as the
author is aware), 9 are considered partially solved, three of them are unsolved and two of them are
considered too vague. Unsolved problems include the infamous Riemann Hypothesis and an extension to
the Kronecker-Weber Theorem. Amongst solved problems is the 10th Hilbert problem.

@@

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} vector sorting}
  \begin{algorithmic}[1]
    \Require\, $n\in\mathbb{N}$ and a vector $A[1..n]$ with $n$ integers
    \Ensure\, An ordered vector
    \Description\, Sorting the vector $A$ in increasing order.
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Search problem:} greatest common divisor}
  \begin{algorithmic}[1]
    \Require\,$a, b\neq 0\in\mathbb{N}$
    \Ensure\, $d|a$, $d|b$, and for all $d'|a$, $d'|b$ we have $d'\leq d$
    \Description\, Finding the $\gcd(a, b)$
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Optimization problem:} longest common subsequence}
  \begin{algorithmic}[1]
    \Require\, Two strings $X[1..m]$ and $Y[1..n]$
    \Ensure\, A string
    \Description\, Finding the longest common subsequence in $X$ and $Y$.
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Decision problem:} greatest common divisor}
  \begin{algorithmic}[1]
    \Require\,$a, b\neq 0\in\mathbb{N}$ and $k\in\mathbb{N}$
    \Ensure\, Boolean value
    \Description\, Is $\gcd(a, b)=k$ true?
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
  \caption*{\textbf{Decision problem:} longest common subsequence}
  \begin{algorithmic}[1]
    \Require\, Two strings $X[1..m]$ and $Y[1..n]$
    \Ensure\, Boolean value
    \Description\, Is there an LCS of X and Y such that its length $\geq k$?
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} Hamiltonian cycle}
  \begin{algorithmic}[1]
    \Require\, A graph $G=(V, E)$
    \Ensure\, Boolean value
    \Description\, Is there a Hamiltonian cycle, i.e.\ a path $p=(e_1,e_2,\ldots,e_k)$ s.t. $\forall v
    \in V$, $v$ is visited by $p$ exactly once, and $p$ is a cycle?
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} SAT}
  \begin{algorithmic}[1]
    \Require\, A Boolean formula with $n$ variables written in CNF, e.g.\ $(x_1\vee x_2\vee x_2\vee
    x_3)\wedge (\ov{x}_2\vee x_5)\wedge\cdots$
    \Ensure\, Boolean value
    \Description\, Is there a valuation $\nu :\{x_1,x_2,\ldots,x_n\}\to\{\false, \true\}$ such that the
    input formula evaluates to true?
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} 3-SAT}
  \begin{algorithmic}[1]
    \Require\, A Boolean formula with $n$ variables written in CNF with at most 3 literals in each
    clause, e.g.\ $(x_1\vee x_2\vee x_3)\wedge(\ov{x}_2\vee\ov{x}_1\vee x_5) \wedge\cdots$
    \Ensure\, Boolean value
    \Description\, Is there a valuation $\nu :\{x_1,x_2,\ldots,x_n\}\to\{\false, \true\}$ such that the
    input formula evaluates to true?
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} quadratic congruence}
  \begin{algorithmic}[1]
    \Require\, $\alpha,\beta,\gamma\in\mathbb{Z}$
    \Ensure\, Boolean value
    \Description\, Is there a solution $x\in\mathbb{Z}$ where
    \begin{equation*}
      x^2 \equiv\alpha\mod\beta
    \end{equation*}
    and such that $0\leq x\leq\gamma$?
  \end{algorithmic}
\end{algorithm}

@@

\begin{align*}
  (i) \quad &0\leq x_1\leq H, \quad x_1\in\mathbb{Z},\\
  (ii) \quad &\lambda_1 2\cdot 8^{m+1}(H^2-x_1^2)+\lambda_2 K(\tau^2-x_1^2)\equiv 0\mod 2\cdot
    8^{m+1}\cdot K, &\text{(III)}\\
  (iii) \quad &\gcd(\lambda_1,K)=\gcd(\lambda_2,2\cdot 8^{m+1})=1,\quad\lambda_1,\lambda_2\in
    \mathbb{Z}.
\end{align*}

@@

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} 0\texttt{-}1 knapsack}
  \begin{algorithmic}[1]
    \Require\, Set of $n$ items, weights $w_1,\ldots,w_n$, values $v_1,\ldots,v_n$ and maximum
    capacity $W$, all positive integers
    \Description\, Can a value of at least $V$ be achieved without exceeding $W$?
    \Reduction\, Exact cover
    \Reference\, [Karp, 1972 (\cite{karp})]
  \end{algorithmic}
\end{algorithm}

When describing the algorithm in~\cite{qcp2}, Adleman and Manders commented on the algorithm's
similarity to the Knapsack problem. In fact, we can reduce QCP to Knapsack, as the algorithm for
Knapsack can be used as a subroutine for the algorithm quoted in this paper. In fact, Knapsack is
satisfiable if and only if $\sum_{j=0}^n c_j\alpha_j\equiv\tau\mod 8^{m+1}$.

We end this paper with a list of selected Number Theory problems that were shown to be NP-complete.
This list was extracted from~\cite{garey-johnson}. All the following problems are decision
problems, meaning we will ommit their output, as they should all return a boolean value.  We label
as \textbf{Reduction} the transformation done in the proof (e.g.\ the 3-SAT was reduced to the QCP
in our proof), and as \textbf{Reference} the article with the reduction. The field \textbf{Comment}
contains comments given by Garey and Johnson on each problem~\cite{garey-johnson}.

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} simultaneous incongruences}
  \begin{algorithmic}[1]
    \Require\, Collection $\{(a_1,b_1),\ldots,(a_n,b_n)\}$ of ordered pairs of positive integers,
    with $a_i\leq b_i$ for $1\leq i\leq n$
    \Description\, Is there an integer $x$ such that, for $1\leq i\leq n, x\not\equiv a_i\mod b_i$?
    \Reduction\, 3-SAT
    \Reference\,~\cite{stockmeyer-meyer}
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} root of modulus 1}
  \begin{algorithmic}[1]
    \Require\, Ordered pairs $(a[i],b[i])$, $1\leq i\leq n$, of integers, with each $b[i]\geq 0$.
    \Description\, Does the polynomial $\sum_{i=1}^n a[i]\cdot z^{b[i]}$ have a root on the complex
    unit circle, i.e., is there a complex number $q$ with $|q|=1$ such that $\sum_{i=1}^n a[i]\cdot
    q^{b[i]}=0$?
    \Reduction\, 3-SAT
    \Reference\,~\cite{plaisted-77b}
    \Commentx\, Not known to be in NP or co-NP\@.
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} cosine product integration}
  \begin{algorithmic}[1]
    \Require\, Sequence $(a_1,a_2,\ldots,a_n)$ of integers
    \Description\, Does $\int_0^{2\pi}(\prod_{i=1}^n \cos(a_i\theta))d\theta=0$?
    \Reduction\, Partition
    \Reference\,~\cite{plaisted-76}
    \Commentx\, Solvable in pseudo-polynomial time.
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} non-trivial greatest common divisor}
  \begin{algorithmic}[1]
    \Require\, Sequences $A_i=<(a_i[1],b_i[1]),\ldots,(a_i[k],b_i[k])>$, $1\leq i\leq m$, of pairs
    of integers, with each $b_i[j]\geq 0$.
    \Description\, Does the $\gcd$ of the polynomials $\sum_{j=1}^k a_i[j]\cdot z^{b_i[j]}$, $1\leq
    i\leq m$, have a degree greater than zero?
    \Reduction\, 3-SAT
    \Reference\,~\cite{plaisted-77a}
    \Commentx\, Not known to be in NP or co-NP\@. Remains NP-hard if each $a_i[j]$ is either $-1$
    or $+1$~\cite{plaisted-76} or if $m=2$~\cite{plaisted-77b}. The analogous problem in which the
    instance also includes a positive integer $K$, and we are asked if the least common multiple of
    the given polynomials has degree less than $K$, is NP-hard under the same restrictions. Both
    problems can be solved in pseudo-polynomial time using standard algorithms.
  \end{algorithmic}
\end{algorithm}
\newpage

\begin{algorithm}[h]
  \caption*{\textbf{Problem:} simultaneous divisibility of linear polynomials}
  \begin{algorithmic}[1]
    \Require\, Vectors $a_i=(a_i[0],\ldots,a_i[m])$ and $b_i=(b_i[0],\ldots,b_i[m])$, $1\leq i\leq
    n$, with positive integer entries
    \Description\, Do there exist positive integers $x_1,x_2,\ldots,x_m$ such that, for $1\leq
    i\leq n$, $a_i[0]+\sum_{j=1}^m(a_i[j]\cdot x_j)$ divides $b_i[0]+\sum_{j=1}^m(b_i[j]\cdot
    x_j)$?
    \Reduction\, Quadratic diophantine equations (i.e.\ \autoref{diophantine-thm})
    \Reference\,~\cite{lipshitz-77,lipshitz-78}
    \Commentx\, Not known to be in NP, but belongs to NP for any fixed $n$. NP-complete for any
    fixed $n\geq 5$. General problem is undecidable if the vector entries and the $x_j$ are allowed
    to range over the ring of ``integers'' in a real quadratic extension of the rationals.
  \end{algorithmic}
\end{algorithm}

%--------------------------------------------------------------------------------------------------
\newpage
